OA Share

### 2020-10-12 analysis/008_subject_exploration.Rmd
update to new datasets

```{r}
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_volume_eligible.csv"))
```

```{r}
ind_per_jn <- hybrid_volume %>%
  group_by(issn, container_title, articles, issued_year) %>%
  summarise(n = n_distinct(doi)) %>%
  mutate(prop = n / articles)
```

@@

```{r}
hybrid_volume %>% 
  filter(!issn %in% scopus_norm$issn) %>% 
  distinct(issn, container_title)
```

@@

```{r}
ajsc_mapped <- read_csv(here::here("data", "asjc_mapped.csv")) %>%
  mutate(top_level_code = as.character(top_level_code))
hybrid_subject <- left_join(ind_per_jn, cite_score_19, by = c("issn" = "issn")) %>%
  mutate(top_level_code = substr(as.character(scopus_asjc_code_sub_subject_area), 1, 2)) %>%
  inner_join(ajsc_mapped, by = "top_level_code")
# back up ind
 write_csv(hybrid_subject, here::here("data", "jn_scopus_ind_subjects.csv"))
# backup subject liste
hybrid_subject %>% 
  ungroup() %>% 
  distinct(issn, container_title, subject_area, top_level = description) %>%
  write_csv(here::here("data", "jn_subjects.csv"))
```

@@

```{r}
hybrid_subject_19 <- hybrid_subject %>% 
  filter(issued_year == "2019", !is.na(percentile)
         #, prop > 0.01 & prop < 0.15) 
         )%>%
  distinct(issn, prop, percentile, snip, sjr)
```

@@

```{r}
hybrid_subject %>%
  filter(issued_year == "2019", !is.na(percentile)) %>% 
  ggplot(aes(prop, snip)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
  facet_wrap(~top_level_code, scales = "free") +
  theme_minimal()
```

@@

```{r}
# sort by median per asjc subject group
subject_all_years <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(n),
            n = sum(articles)) %>%
  ungroup() %>%
  group_by(description) %>%
  filter(subject_area != "Multidisciplinary")

@@

```{r}
funder_shares <- hybrid_volume %>%
   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  group_by(oa_sponsor_type, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_volume_jn, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
  left_join(jn_subjects, by = "issn") %>%
  ungroup() %>%
  group_by(top_level, oa_sponsor_type) %>%
  summarise(oa_type = sum(n_oa_type)) %>%
  mutate(prop = oa_type / sum(oa_type))
# overall props
oa_sponsor_type_all <- hybrid_volume %>% 
  #exclude multidisciplinary
  filter(issn != "2095-9273") %>%
  count(oa_sponsor_type) %>% 
  mutate(prop = n / sum(n))

@@

```{r}
field_funder <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>% 
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
   mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  left_join(jn_subjects, by = "issn") %>%
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  group_by(oa_sponsor, top_level) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
  mutate(oa_sponsor = fct_reorder(oa_sponsor, n_oa_type))
ggplot(field_funder, aes(x = oa_sponsor, 
                         y = reorder(top_level, desc(top_level)), 
                         colour = prop, size = n_oa_type)) + 
  geom_point() +
  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
                      option = "magma",
                      name = "Percentage per funder",
                      breaks =  scales::extended_breaks()) +
  scale_x_discrete(position = "top") +
  labs(x = NULL, y = NULL) +
  theme_minimal_grid() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank()) +
  theme(legend.position = "top", legend.justification = "right") +
  theme(legend.direction = "horizontal", legend.box = "vertical") +
  theme(strip.text = element_text(face="bold"))
```

>>>>

```{r}
hybrid_oa_volume <- readr::read_csv(here::here("data", "hybrid_oa_volume.csv"))
```

@@

```{r}
hybrid_oa_volume %>% 
  filter(!issn %in% scopus_norm$issn) %>% 
  distinct(issn)
```

@@

```{r}
ajsc_mapped <- read_csv(here::here("data", "asjc_mapped.csv")) %>%
  mutate(top_level_code = as.character(top_level_code))
hybrid_subject <- left_join(hybrid_oa_volume, cite_score_19, by = c("issn" = "issn")) %>%
  mutate(top_level_code = substr(as.character(scopus_asjc_code_sub_subject_area), 1, 2)) %>%
  inner_join(ajsc_mapped, by = "top_level_code")
# back up ind
 write_csv(hybrid_subject, here::here("data", "jn_scopus_ind_subjects.csv"))
# backup subject liste
hybrid_subject %>% 
  ungroup() %>% 
  distinct(issn, subject_area, top_level = description) %>%
  write_csv(here::here("data", "jn_subjects.csv"))
```

@@

```{r}
hybrid_subject_19 <- hybrid_subject %>% 
  mutate(prop = oa /articles) %>%
  filter(issued_year == "2019", !is.na(percentile)
         #, prop > 0.01 & prop < 0.15) 
         )%>%
  distinct(issn, prop, percentile, snip, sjr)
```

@@

```{r}
hybrid_subject %>%
  mutate(prop = oa /articles) %>%
  filter(issued_year == "2019", !is.na(percentile)) %>% 
  ggplot(aes(prop, snip)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
  facet_wrap(~top_level_code, scales = "free") +
  theme_minimal()
```

@@

```{r}
# sort by median per asjc subject group
subject_all_years <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  ungroup() %>%
  group_by(description) %>%
  filter(subject_area != "Multidisciplinary")

@@

```{r}
hybrid_jn_volume <- hybrid_oa_volume %>%
  group_by(issn) %>% 
  summarise(n_oa = sum(oa))

hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
funder_shares <- hybrid_df %>%
   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  group_by(oa_sponsor_type, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_jn_volume, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
  left_join(jn_subjects, by = "issn") %>%
  ungroup() %>%
  group_by(top_level, oa_sponsor_type) %>%
  summarise(oa_type = sum(n_oa_type)) %>%
  mutate(prop = oa_type / sum(oa_type))
# overall props
oa_sponsor_type_all <- hybrid_df %>% 
  #exclude multidisciplinary
  filter(issn != "2095-9273") %>%
  count(oa_sponsor_type) %>% 
  mutate(prop = n / sum(n))

@@

```{r}
field_funder <- hybrid_df %>%
  filter(oa_sponsor_type == "FundingBody") %>% 
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
   mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  left_join(jn_subjects, by = "issn") %>%
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  group_by(oa_sponsor, top_level) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
  mutate(oa_sponsor = fct_reorder(oa_sponsor, n_oa_type))
ggplot(field_funder, aes(x = oa_sponsor, 
                         y = reorder(top_level, desc(top_level)), 
                         colour = prop, size = n_oa_type)) + 
  geom_point() +
  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
                      option = "magma",
                      name = "Percentage per funder",
                      breaks =  scales::extended_breaks()) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 6), position = "top") +
  labs(x = NULL, y = NULL) +
  theme_minimal_grid(font_size = 10) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank()) +
  theme(legend.position = "top", legend.justification = "right") +
  theme(legend.direction = "horizontal", legend.box = "vertical") 
```
